---
title: "Summary Report of KMeans on Holiday Dataset"
author: | 
    | Team B :
    |     Felicia, Cazer --------------A00225804
    |     Anup, Kumar Singh ------A00258773
    |     Achal, Gupta ---------------A00258772
    |     Toro, Keng Yee Onn------A00214412
date: "5 April 2019"
output: html_document
---
#Data Set Information:
This dataset was populated from destination reviews published by 249 reviewers of holidayiq.com till October 2014. Reviews falling in 6 categories among destinations across South India were considered and the count of reviews in each category for every reviewer (traveler) is captured. <br>

#Attribute Information
Attribute 1 : Unique user id <br>
Attribute 2 : Number of reviews on stadiums, sports complex, etc. <br>
Attribute 3 : Number of reviews on religious institutions <br>
Attribute 4 : Number of reviews on beach, lake, river, etc. <br>
Attribute 5 : Number of reviews on theatres, exhibitions, etc. <br>
Attribute 6 : Number of reviews on malls, shopping places, etc. <br>
Attribute 7 : Number of reviews on parks, picnic spots, etc.<br>


```{r include=FALSE, warning = FALSE, message = FALSE}
###IMPORTING LIBRARIES
#Below code will import all the required libraries.
#install.packages("NbClust")
#install.packages("factoextra")
#install.packages("psych")
#installed.packages("ggplot2")
#install.packages("dplyr")

library(NbClust)
library(dplyr)
library(psych)
library(factoextra) 
library(ggplot2)
```

```{r include=FALSE, warning = FALSE}
###DATA ACCQUISTION
#Below code will import the data present in CSV format into R and store it in a Data Frame.
Holiday_Data <- read.csv("Data/buddymove_holidayiq.csv")

#View the data
str(Holiday_Data,5)
```

```{r include=FALSE, warning = FALSE}
###DATA CLEANING
#Omit rows with Null values if any from the data
Holiday_Data <- na.omit(Holiday_Data)
```

```{r include=FALSE, warning = FALSE, tidy = TRUE}
#### Creating subset of required columns
#Create a subset without User Id column as for K means only numerical columns are chosen 
Holiday_Data_num <- Holiday_Data[,-1]
```


#### Scaling data
Visualizing data distribution before scaling using boxplot
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
boxplot(Holiday_Data_num, outline= FALSE)
```
</br>Since clustering algorithms including kmeans use distance-based measurements to determine the similarity between data points, it’s recommended to standardize the data to have a mean of zero and a standard deviation of one since almost always the features in any dataset would have different units of measurements.
```{r include=FALSE, warning = FALSE, tidy = TRUE}
Holiday_Data_num <- as.data.frame(lapply(Holiday_Data_num, scale))
```
Visualizing data distribution after scaling
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
boxplot(Holiday_Data_num, outline= FALSE)

```
</br>
```{r include=FALSE, warning = FALSE, tidy = TRUE}
#verify mean = 0 and standard deviation = 1
describe(Holiday_Data_num, fast=T)
```

```{r include=FALSE, warning = FALSE, tidy = TRUE}
#assigning background color and border line for graphs
fill <- "gold1"
line <- "goldenrod2"
```

### Descriptive Analysis
```{r include=FALSE, warning = FALSE, tidy = TRUE}
#Duplicate data just for exploatory purpose, we will be creating some calculated columns and don't want to alter main dataset.
Holiday_Data_Explore <- Holiday_Data
```

#### Top 5 active users based on total reviews given
```{r include=FALSE, warning = FALSE, tidy = TRUE}
#Identify Total reviews of each user
Holiday_Data_Explore$Total <- rowSums(Holiday_Data_Explore[,c("Sports","Religious","Nature", "Theatre", "Shopping","Picnic")]) 
```

```{r include=FALSE, warning = FALSE, tidy = TRUE}
#Sort data based on Total Number of reviews
Holiday_Data_Explore <- Holiday_Data_Explore %>% 
  arrange(Total)
```
Visualize Top 5 active users using bar graph
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
ggplot(data=tail(Holiday_Data_Explore,5), aes(x=User.Id, y=Total)) +
  geom_bar(stat = "identity",fill = fill, colour = line) +
  geom_text(aes(label=Total), vjust=1.6, color="black", size=3.5)+
  theme_minimal() + 
  ggtitle("Top 5 Active Users") + 
  xlab("User Id") +
  ylab("Total NUmber of Reviews")
```

#### Top 5 shopaholics users
```{r include=FALSE, warning = FALSE, tidy = TRUE}
#Sort data based on Total Number of shopping reviews
Holiday_Data_Explore <- Holiday_Data_Explore %>% 
  arrange(Shopping)
```
Bar graph to repreent users who shop most
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
ggplot(data=tail(Holiday_Data_Explore,5), aes(x=User.Id, y=Total)) +
  geom_bar(stat = "identity",fill = fill, colour = line) +
  geom_text(aes(label=Total), vjust=1.6, color="black", size=3.5)+
  theme_minimal() + 
  ggtitle("Top 5 Shopaholics") + 
  xlab("User Id") +
  ylab("Total NUmber of Reviews")

```

Graph below will perform three analysis</br>
1. Distribution of data for each feature or column shown using a histogram.</br>
2. Correlation coefficients for each combination of feature</br>
3. Scatterplot with fit line to visualize correlation</br>
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
pairs.panels(Holiday_Data_Explore[,c("Sports","Religious","Nature", "Theatre", "Shopping","Picnic")])
```
</br><b> Key Observations </b>
1. Number of shopping reviews has a strong positive relation with number of relegious destination reviews.</br>
2. Number of reviews by user for Shopping and Nature destinations shows weak negative corelation </br>
3. Number of reviews by user for Relegious and Nature destinations shows weak negative corelation </br>
4. Number of reviews by user for Nature and Theatre destinations shows weak positive corelation.</br>


#### Most Reviewed Destinations

```{r include=FALSE, warning = FALSE, tidy = TRUE}
#First get column wise sums for each destination type and then sort it in decreaseing order
mostVoted <- order(colSums(Holiday_Data_Explore[,c("Sports","Religious","Nature", "Theatre", "Shopping","Picnic")]),decreasing=TRUE)
```

```{r include = FALSE, warning = FALSE, tidy = TRUE}
#Get the name of the column from main data frame
#Note <- +1 is done because in mostVoted we ignored first column i.e. User.Id
colnames(Holiday_Data_Explore)[mostVoted[1]+1]
```

#### K Means Clustering
##### Identify best value of K

```{r include=FALSE, warning = FALSE, tidy = TRUE}
#create empty lists
k <- list()
betweenss <- list()
totalss <- list()
ratio <- list()
```

```{r include=FALSE, warning = FALSE, tidy = TRUE}
#Elbow method to identify K value.
#For loop will calculate ratio of between cluster sum of squares and Total within clusters sum of squares for clusters of k value 1 to 15
for(i in 1:15){
  k[[i]] <- kmeans(Holiday_Data_num, i, iter.max = 100)
  betweenss[[i]] <- k[[i]]$betweenss
  totalss[[i]] <- k[[i]]$tot.withinss
  ratio[[i]] <- k[[i]]$betweenss/k[[i]]$totss
}
plot(1:15, ratio, type = 'b', ylab = 'Between SS / Total SS', xlab = "Clusters(k)")
```

```{r warning=FALSE, echo=FALSE, tidy=TRUE}
#either 2 or 3 is the best value

#Using nbclust package to verify same
set.seed(1234)
nc <- NbClust(Holiday_Data_num, min.nc=2, max.nc=15, method="kmeans")
```
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
table(nc$Best.n[1,])
```
K= 2 and 3 both shows maximum values, geneally an odd value for k is preffered and hence k=3 is optimumu value.

Let's make model using both values and evaluate.
```{r include=FALSE, warning = FALSE, tidy = TRUE}
cluster_up <- kmeans(Holiday_Data_num, 2, iter.max = 100, nstart = 25)
```
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
fviz_cluster(cluster_up, data = Holiday_Data_num)
```
</br>
Table of average of various columns for each cluster
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
aggregate(Holiday_Data[-1], by=list(cluster=cluster_up$cluster), mean)
```
Not much insight can be gathered from 2 clusters it just divide data into two categories user with less reviews on all groups i.e. new users and people with more reviews i.e outgoing or active users

Lets crete cluster with more than 2 groups i.e. Building cluster with k=3
```{r include=FALSE, warning = FALSE, tidy = TRUE}
cluster_up <- kmeans(Holiday_Data_num, 3, iter.max = 100, nstart = 25)
```

Analysing the cluster
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
print(cluster_up)
```

Above output shows the sizes of three clusters formed along with the mean value for each column for each cluster.</br> 
Then the output displays a vector of cluster value for each row.</br>
Within cluster sum of squares is displayed for each of the three clusters.</br>

```{r echo=FALSE, warning = FALSE, tidy = TRUE}
str(cluster_up)
```
Above output displays following information</br>
<b>Cluster: </b>A vector of integers (from 1:k) indicating the cluster to which each point is allocated. </br>
<b>Centers: </b>A matrix of cluster centers. </br>
<b>totss: </b>The total sum of squares. </br>
<b>withinss: </b>Vector of within-cluster sum of squares, one component per cluster. </br>
<b>tot.withinss: </b>Total within-cluster sum of squares, i.e. sum(withinss). </br>
<b>betweenss: </b>The between-cluster sum of squares, i.e. $totss-tot.withinss$. </br>
<b>size: </b>The number of points in each cluster </br>

```{r echo=FALSE, warning = FALSE, tidy = TRUE}
fviz_cluster(cluster_up, data = Holiday_Data_num)
```
</br>Creating table of average of various columns for each cluster
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
aggregate(Holiday_Data[-1], by=list(cluster=cluster_up$cluster), mean)
```
Below traits about useers can be assumed from the cluster information above.
Cluster 3 - New Users with less reviews
Cluster 2 - Young Artistic Crowd - More into Nature, Theatre
Cluster 1 - Family oriented person - More into Religious institutions, shopping

Visualizing clusters for various combination of scatterplots
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
plot(Holiday_Data_num, col=cluster_up$cluster)
```

Focusing on some major insights:
Assuming less number of reviews by a user on a particular destination means user visit that destination group less number of times.
People who visit more relegious institutions tends to do more shopping reviews as can be viewed from the plot below
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
plot(Holiday_Data[c("Shopping","Religious")], col=cluster_up$cluster, 
     main = "Number of User reviews for Shopping vs Relegious destinations")
```
</br>People who are nature lovers do less shopping as can be viewed from the plot below
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
plot(Holiday_Data[c("Shopping","Nature")], col=cluster_up$cluster,
     main = "Number of User reviews for Shopping vs Nature destinations")
```
</br>People who are nature lovers are more into Theatre destinations as can be viewed from the plot below
```{r echo=FALSE, warning = FALSE, tidy = TRUE}
plot(Holiday_Data[c("Theatre","Nature")], col=cluster_up$cluster,
     main = "Number of User reviews for Theatre vs Nature destinations")
```